Staged Abstract Interpreters

### 2019-02-20 paper/10_intro.tex
revision

In this paper, we study the confluence of two old ideas --- Futamura projection
and abstract interpretation, but from the perspective of their recent
realizations --- multi-stage programming and definitional abstract interpreters.
To be specific, we present the application of the first Futamura projection on
definitional abstract interpreters, which intersects as staged abstract
interpreters. By borrowing the ideas from \note{monadic interpreters}
\cite{DBLP:conf/popl/LiangHJ95, DBLP:journals/pacmpl/DaraisLNH17,
Sergey:2013:MAI:2491956.2491979} and embedded domain-specific languages
\cite{DBLP:conf/snapl/RompfBLSJAOSKDK15, DBLP:journals/jfp/CaretteKS09,
DBLP:conf/icfp/GibbonsW14}, we develop an approach to fulfill four different
semantics (Figure~\ref{confluence}) in a unified framework. The four semantics
share one same generic interpreter and some abstract types; by instantiating the
abstract types differently, the interpreter behaves as a concrete interpreter,
an abstracter interperter, or their staged versions, respectively.

% FIXME: and the approach of constructing optimizing abstract interpreters by multi-stage programming.

>>>>

In this paper, we study the confluence of two old ideas --- Futamura projection
and abstract interpretation, but from the perspective of their recent
realizations --- multi-stage programming and definitional abstract interpreters.
To be specific, we present the application of the first Futamura projection on
definitional abstract interpreters, which intersects as staged abstract
interpreters. By borrowing the ideas from \note{monadic interpreters}
\cite{DBLP:conf/popl/LiangHJ95, DBLP:journals/pacmpl/DaraisLNH17,
Sergey:2013:MAI:2491956.2491979} and embedded domain-specific languages
\cite{DBLP:conf/snapl/RompfBLSJAOSKDK15, DBLP:journals/jfp/CaretteKS09,
DBLP:conf/icfp/GibbonsW14}, we develop an approach to fulfill four different
semantics (Figure~\ref{confluence}) in a unified framework. The four semantics
share one same generic interpreter. However, by instantiating the interpreter
differently, it can behave as a concrete interpreter, an abstracter interperter,
or their staged versions, respectively. The key idea to enable this flexibility
is to use abstract type members \note{or type parameters} abstract over concrete/abstract components
(e.g., concrete values or abstract values) of the interpreter, as well as the
binding-time (e.g., static or dyanmic).

### 2019-02-20 paper/20_bg.tex
revision

\section{Background}

In this section, we first described the language we will model and analyze, and present an unstaged concrete 
interpreter for it. Then, starting from this unstaged concrete interpreter, we proceed in two directions:
by abstracting it we obtain an unstaged abstract interpreter; by staging it we obtain a staged concrete
interpreter.

>>>>

\section{Preliminaries}

In this section, we first describe the abstract syntax of the language we will
analyze, then present the generic interface shared between the four different
interpreters. We first instantiate the interpreter to the concrete one; then we
proceed in two directions: by staging it we obtain a staged concrete interpreter
(Section~\ref{stagedinterp}), by abstracting it we obtain an unstaged abstract
interpreter (Section~\ref{unstaged_abs});

### 2019-02-20 paper/21_bg_interp.tex
intro
revision

We consider a tiny higher-order functional language \TLang, which is based on
direct-style $\lambda$-calculus with numbers, arithmetics, recursions and
conditionals. In Section~\ref{cases_imp}, we will add more imperative features
to the language. Since we are mostly interested in analyzing the dynamic
behavior of the program, we disguised any static semantics and type system. We
also assume that input programs are well-typed and variables are distinct. The
abstract syntax is as follows:

@@

\subsection{Generic Interface}

>>>>

We consider a call-by-value $\lambda$-calculus in direct-style, but extended
with numbers, arithmetics, recursions and conditionals. Other effectful features
such as assignments also can be supported easily; in Section~\ref{cases_imp}, we
will add more imperative features to the language. Since we are mostly
interested in analyzing the dynamic behaviors of the program, we disguised any
static semantics and type system. We also assume that input programs are
well-typed and all variables are distinct. The abstract syntax is shown as
follows:

@@

\subsection{Generic Interface} \label{generic_if}
